---
title: "HAARVI COVID-19: Assign Patients to Batches"
author: "Malisa Smith"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
date: "version `r format(Sys.time(), '%B %d, %Y')`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F}
library(tidyverse)
library(ggpubr)
```

```{r}
save_output <- FALSE
```

# Read in data

```{r read in data}
# Read in the patient manifest
data_manifest <- read.csv(here::here("data/Seshadri_HAARVI_PBMC_manifest_merged_02Oct2020.csv"), check.names = F, stringsAsFactors = F)
unique(data_manifest$Cohort)
```
# Sample and assign patients to batch

Samples to be run:  
(1) Convalescent donors that were previously hospitalized (n=20)  
(2) Convalescent donors that were never hospitalized (n=40)  
(3) Healthy controls from Chu lab (n=4) and TRIMAS (n=4, 3 from Seshadri and 1 from Kiel)  
Seshadri TRIMAS: BWT22, BWT23, BWT20  
  
Split into 3 batches (1 batch of 22 and 2 batches of 23):  
Batch 1 (n=22): ConvHosp=6, ConvNotHosp=14, Healthy=1, TRIMA=1  
Batch 2 (n=23): ConvHosp=7, ConvNotHosp=13, Healthy=1, TRIMA=2  
Batch 3 (n=23): ConvHosp=7, ConvNotHosp=13, Healthy=2, TRIMA=1  

## First prepare data for sampling

```{r}
# Add TRIMA rows and prepare data for sampling
all_data <- data_manifest %>% 
  bind_rows(data.frame(`Record ID` = c("BWT20", "BWT22", "BWT23", "ChuLab_TRIMA_1"),
                       Cohort = "TRIMA",
                       check.names = F)) %>% 
  mutate(Classification = dplyr::recode(Cohort,
                                   "Non-hospitalized" = "ConvNotHosp",
                                   "Hospitalized" = "ConvHosp",
                                   "Healthy control 2017-2018" = "Healthy",
                                   "Healthy control" = "Healthy")) %>% 
  mutate(Classification = factor(Classification, levels = c("ConvHosp", "ConvNotHosp", "Healthy", "TRIMA")))

length(unique(all_data$`Record ID`)) == nrow(all_data) # Confirm each Record ID is unique

knitr::kable(head(all_data))
```

## Sample

```{r}
set.seed(20200522)

# Sample for Batch 1
batch1 <- all_data %>%
  group_by(Classification) %>% # prep for work by group
  nest() %>% # one row per group           
  ungroup() %>% 
  arrange(Classification) %>% 
  mutate(n = c(6, 14, 1, 1)) %>% # add sample sizes
  mutate(samp = map2(data, n, sample_n)) %>% 
  select(-data, -n) %>%
  unnest(samp) %>%
  arrange(Classification, `Record ID`)

cat("Batch 1 Patients:")
knitr::kable(batch1 %>% 
               dplyr::select(Classification, `Record ID`))

# Sample for Batch 2
batch2 <- all_data %>% 
  anti_join(batch1, by="Record ID") %>% 
  group_by(Classification) %>% # prep for work by group
  nest() %>% # one row per group           
  ungroup() %>% 
  arrange(Classification) %>% 
  mutate(n = c(7, 13, 1, 2)) %>% # add sample sizes
  mutate(samp = map2(data, n, sample_n)) %>% 
  select(-data, -n) %>%
  unnest(samp) %>%
  arrange(Classification, `Record ID`)

cat("Batch 2 Patients:")
knitr::kable(batch2 %>% 
               dplyr::select(Classification, `Record ID`))

# Batch 3
batch3 <- all_data %>% 
  anti_join(batch1, by="Record ID") %>% 
  anti_join(batch2, by="Record ID") %>% 
  arrange(Classification, `Record ID`)

cat("Batch 3 Patients:")
knitr::kable(batch3 %>% 
               dplyr::select(Classification, `Record ID`))
```

```{r}
# Confirm that there are correct numbers of each group in each batch
batch1 %>% group_by(Classification) %>% summarize(PTIDs_in_group = n_distinct(`Record ID`)) %>% knitr::kable()
batch2 %>% group_by(Classification) %>% summarize(PTIDs_in_group = n_distinct(`Record ID`)) %>% knitr::kable()
batch3 %>% group_by(Classification) %>% summarize(PTIDs_in_group = n_distinct(`Record ID`)) %>% knitr::kable()

intersect(batch1$`Record ID`, batch2$`Record ID`) # And none of them overlap
intersect(batch1$`Record ID`, batch3$`Record ID`)
intersect(batch2$`Record ID`, batch3$`Record ID`)

if(save_output) {
  write.csv(batch1, here::here("processed_data/20200522_COVID_ICS_DURT_Batch_1.csv"), row.names = FALSE)
  write.csv(batch2, here::here("processed_data/20200522_COVID_ICS_DURT_Batch_2.csv"), row.names = FALSE)
  write.csv(batch3, here::here("processed_data/20200522_COVID_ICS_DURT_Batch_3.csv"), row.names = FALSE)
}

```