---
title: "COVID ICS & DURT: Assign Patients to Batches"
author: "Malisa Smith"
date: "May 22, 2020"
output:
  rmdformats::html_clean:
    highlight: kate
  toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F}
library(readxl)
library(tidyverse)
library(ggpubr)
```

```{r}
save_output <- FALSE
```

# Read in data

```{r read in data}
# Read in the patient manifest
data_chu <- readxl::read_excel(here::here("data/Chu_Seshadri_HAARVI_Collab_PBMC_manifest_10June2020.xlsx"))
unique(data_chu$Cohort)
```

# Check if matched by age, sex, ethnicity, and date of symptom onset

```{r}
# First explore missing data
data_chu %>% 
  dplyr::filter(Sex == "N/A" | Race == "N/A" | `Days symptom onset to visit 1` == "N|A" | Age == "N/A") %>% 
  dplyr::select(`Record ID`, Cohort, Age, Sex, Race, `Days symptom onset to visit 1`) %>% 
  knitr::kable()
```
  
8 hospitalized patients have Sex=N/A and Race=N/A  
No Race information for Healthy Controls  
No "Days symptom onset to visit 1" information for 2 hopitalized patients (not relevant for healthy controls)    

```{r, fig.width=5.5, fig.height=4}
# Age
data_chu %>% 
  dplyr::filter(Cohort %in% c("Non-hospitalized", "Hospitalized")) %>% 
  mutate(Cohort = factor(Cohort, levels = c("Non-hospitalized", "Hospitalized"))) %>% 
  ggboxplot(x="Cohort", y="Age", color="Cohort", add="jitter") +
  theme_grey(base_size=18) +
  stat_compare_means() +
  theme(legend.position = "none")

# Sex
data_chu %>% 
  dplyr::filter(Cohort %in% c("Non-hospitalized", "Hospitalized")) %>% 
  dplyr::filter(Sex != "N/A") %>% 
  mutate(Cohort = factor(Cohort, levels = c("Non-hospitalized", "Hospitalized"))) %>% 
  ggplot(aes(x=Cohort, fill=Sex)) +
  theme_grey(base_size=18) +
  geom_bar(position="fill", stat="count")
sex_table <- data_chu %>% 
  dplyr::filter(Cohort %in% c("Non-hospitalized", "Hospitalized")) %>% 
  dplyr::filter(Sex != "N/A") %>% 
  mutate(Cohort = factor(Cohort, levels = c("Non-hospitalized", "Hospitalized"))) %>%
  dplyr::select(Cohort, Sex) %>% table()
sex_table
fisher.test(sex_table)

data_chu %>% 
  dplyr::filter(Cohort %in% c("Non-hospitalized", "Hospitalized")) %>% 
  mutate(Cohort = factor(Cohort, levels = c("Non-hospitalized", "Hospitalized"))) %>% 
  ggplot(aes(x=Cohort, fill=Sex)) +
  theme_grey(base_size=18) +
  geom_bar(position="fill", stat="count")
```

```{r, fig.width=8.5, fig.height=4}
# Ethnicity
data_chu %>% 
  dplyr::filter(Cohort %in% c("Non-hospitalized", "Hospitalized")) %>% 
  dplyr::filter(Race != "N/A") %>% 
  mutate(Cohort = factor(Cohort, levels = c("Non-hospitalized", "Hospitalized"))) %>% 
  mutate(Race_v2 = dplyr::recode(Race,
                                 "White,Prefer not to say"="White",
                                 "American Indian or Alaska Native,White"="American Indian or Alaska Native",
                                 "white"="White",
                                 "americanIndianOrAlaskaNative"="American Indian or Alaska Native",
                                 "other|other"="Other")) %>% 
  ggplot(aes(x=Cohort, fill=Race_v2)) +
  theme_grey(base_size=18) +
  geom_bar(position="fill", stat="count")
race_table <- data_chu %>% 
  dplyr::filter(Cohort %in% c("Non-hospitalized", "Hospitalized")) %>% 
  dplyr::filter(Race != "N/A") %>% 
  mutate(Cohort = factor(Cohort, levels = c("Non-hospitalized", "Hospitalized"))) %>% 
  mutate(Race_v2 = dplyr::recode(Race,
                                 "White,Prefer not to say"="White",
                                 "American Indian or Alaska Native,White"="American Indian or Alaska Native",
                                 "white"="White",
                                 "americanIndianOrAlaskaNative"="American Indian or Alaska Native",
                                 "other|other"="Other")) %>% 
  dplyr::select(Cohort, Race_v2) %>% 
  table()
race_table
fisher.test(race_table)

data_chu %>% 
  dplyr::filter(Cohort %in% c("Non-hospitalized", "Hospitalized")) %>% 
  mutate(Cohort = factor(Cohort, levels = c("Non-hospitalized", "Hospitalized"))) %>% 
  mutate(Race_v2 = dplyr::recode(Race,
                                 "White,Prefer not to say"="White",
                                 "American Indian or Alaska Native,White"="American Indian or Alaska Native",
                                 "white"="White",
                                 "americanIndianOrAlaskaNative"="American Indian or Alaska Native",
                                 "other|other"="Other")) %>% 
  ggplot(aes(x=Cohort, fill=Race_v2)) +
  theme_grey(base_size=18) +
  geom_bar(position="fill", stat="count")
```

```{r, fig.width=5.5, fig.height=4}
# Days from symptom onset to collection
data_chu %>% 
  dplyr::filter(Cohort %in% c("Non-hospitalized", "Hospitalized")) %>% 
  mutate(Cohort = factor(Cohort, levels = c("Non-hospitalized", "Hospitalized"))) %>% 
  dplyr::filter(`Days symptom onset to visit 1` != "N/A") %>% 
  mutate(`Days symptom onset to visit 1` = as.numeric(`Days symptom onset to visit 1`)) %>% 
  ggboxplot(x="Cohort", y="`Days symptom onset to visit 1`", color="Cohort", add="jitter") +
  theme_grey(base_size=18) +
  stat_compare_means() +
  theme(legend.position = "none")
```

# Sample and assign patients to batch

Samples to be run:  
(1) Convalescent donors that were previously hospitalized (n=20)  
(2) Convalescent donors that were never hospitalized (n=40)  
(3) Healthy controls from Chu lab (n=4) and TRIMAS (n=4, 3 from Seshadri and 1 from Kiel)  
Seshadri TRIMAS: BWT22, BWT23, BWT20  
  
Split into 3 batches (1 batch of 22 and 2 batches of 23):  
Batch 1 (n=22): ConvHosp=6, ConvNotHosp=14, Healthy=1, TRIMA=1  
Batch 2 (n=23): ConvHosp=7, ConvNotHosp=13, Healthy=1, TRIMA=2  
Batch 3 (n=23): ConvHosp=7, ConvNotHosp=13, Healthy=2, TRIMA=1  

## First prepare data for sampling

```{r}
# Add TRIMA rows and prepare data for sampling
all_data <- data_chu %>% 
  dplyr::filter(Cohort != "influenza 2018-2019") %>% 
  bind_rows(data.frame(`Record ID` = c("BWT20", "BWT22", "BWT23", "ChuLab_TRIMA_1"),
                       Cohort = "TRIMA",
                       check.names = F)) %>% 
  mutate(Classification = dplyr::recode(Cohort,
                                   "Non-hospitalized" = "ConvNotHosp",
                                   "Hospitalized" = "ConvHosp",
                                   "Healthy control 2017-2018" = "Healthy",
                                   "Healthy control" = "Healthy")) %>% 
  mutate(Classification = factor(Classification, levels = c("ConvHosp", "ConvNotHosp", "Healthy", "TRIMA")))

length(unique(all_data$`Record ID`)) == nrow(all_data) # Confirm each Record ID is unique

knitr::kable(head(all_data))
```

## Use nesting and sampling to sample each batch (help from [here](https://jennybc.github.io/purrr-tutorial/ls12_different-sized-samples.html):

```{r}
set.seed(20200522)

# Sample for Batch 1
batch1 <- all_data %>%
  group_by(Classification) %>% # prep for work by group
  nest() %>% # one row per group           
  ungroup() %>% 
  arrange(Classification) %>% 
  mutate(n = c(6, 14, 1, 1)) %>% # add sample sizes
  mutate(samp = map2(data, n, sample_n)) %>% 
  select(-data, -n) %>%
  unnest(samp) %>%
  arrange(Classification, `Record ID`)

cat("Batch 1 Patients:")
knitr::kable(batch1 %>% 
               dplyr::select(Classification, `Record ID`))

# Sample for Batch 2
batch2 <- all_data %>% 
  anti_join(batch1, by="Record ID") %>% 
  group_by(Classification) %>% # prep for work by group
  nest() %>% # one row per group           
  ungroup() %>% 
  arrange(Classification) %>% 
  mutate(n = c(7, 13, 1, 2)) %>% # add sample sizes
  mutate(samp = map2(data, n, sample_n)) %>% 
  select(-data, -n) %>%
  unnest(samp) %>%
  arrange(Classification, `Record ID`)

cat("Batch 2 Patients:")
knitr::kable(batch2 %>% 
               dplyr::select(Classification, `Record ID`))

# Batch 3
batch3 <- all_data %>% 
  anti_join(batch1, by="Record ID") %>% 
  anti_join(batch2, by="Record ID") %>% 
  arrange(Classification, `Record ID`)

cat("Batch 3 Patients:")
knitr::kable(batch3 %>% 
               dplyr::select(Classification, `Record ID`))
```

```{r}
# Confirm that there are correct numbers of each group in each batch
batch1 %>% group_by(Classification) %>% summarize(PTIDs_in_group = n_distinct(`Record ID`)) %>% knitr::kable()
batch2 %>% group_by(Classification) %>% summarize(PTIDs_in_group = n_distinct(`Record ID`)) %>% knitr::kable()
batch3 %>% group_by(Classification) %>% summarize(PTIDs_in_group = n_distinct(`Record ID`)) %>% knitr::kable()

intersect(batch1$`Record ID`, batch2$`Record ID`) # And none of them overlap
intersect(batch1$`Record ID`, batch3$`Record ID`)
intersect(batch2$`Record ID`, batch3$`Record ID`)

if(save_output) {
  write.csv(batch1, here::here("processed_data/20200522_COVID_ICS_DURT_Batch_1.csv"), row.names = FALSE)
  write.csv(batch2, here::here("processed_data/20200522_COVID_ICS_DURT_Batch_2.csv"), row.names = FALSE)
  write.csv(batch3, here::here("processed_data/20200522_COVID_ICS_DURT_Batch_3.csv"), row.names = FALSE)
}

```